We choose Magadino and Bern as localities.

Import libraries and define options

# install.packages("rmarkdown")

```{r}
library(tidyverse)
library(chron)
```

```{r}
# source("functions_extra.R")
```


```{r}
Sys.setlocale("LC_TIME","C")
```

```{r}
options(stringsAsFactors=FALSE)
options(chron.year.abb=FALSE)
theme_set(theme_bw()) # preferences for plots
```


Start working with data

```{r}
filename1 <- file.path("Data", "MAG.csv")
filename2 <- file.path("Data", "BER.csv")
file.exists(filename1)
file.exists(filename2)
```


```{r}
MAG <- read.table(filename1,sep=";",skip=6,
  col.names=c("datetime","O3","NO2","SO2","PM10","PM2.5", "EC", "NOX", "TEMP","PREC","RAD"))

BER <- read.table(filename2,sep=";",skip=6,
  col.names=c("datetime","O3","NO2","CO","PM10","PM2.5", "EC", "CPC", "NOX", "TEMP","PREC","RAD"))
```

```{r}
head(MAG)
```

```{r}
head(BER)
```

```{r}
str(MAG)
```
```{r}
str(BER)
```

```{r}
MAG[,"datetime"] <- as.chron(MAG[,"datetime"], "%d.%m.%Y %H:%M") - 1/24
MAG[,"month"] <- months(MAG[,"datetime"])
MAG[,"date"] <- dates(MAG[,"datetime"])
```

```{r}
BER[,"datetime"] <- as.chron(BER[,"datetime"], "%d.%m.%Y %H:%M") - 1/24
BER[,"month"] <- months(BER[,"datetime"])
BER[,"date"] <- dates(BER[,"datetime"])
```

Verifiy that there is no missing values or NaN:
```{r}
is.na(MAG)
```

```{r}
is.na(BER)
```
All good, no NaN values to treat!


```{r}
head(MAG)
```

```{r}
head(BER)
```

Check if there is no negative values, which would be unusable: 
```{r}
MAG[MAG < 0, ]
```
```{r}
BER[BER < 0, ]
```
All good, only positive or null values!



First plot of data
For preliminary analysis, we will keep all variables and pollutants for each site.
When we will procede to comparison between the two locations, Bern and Magadino, we will only keep the common pollutants and variables: O3, NO2, PM10, PM2.5, EC, NOX, TEMP, PREC, RAD.

For Magadino: 
View raw ozone concentrations:
```{r}
ggplot(MAG)+
  geom_line(aes(datetime, O3))+
    scale_x_chron()
```

View raw nitrogen dioxide concentrations:
```{r}
ggplot(MAG)+
  geom_line(aes(datetime, NO2))+
    scale_x_chron()
```

View raw sulfur dioxide concentrations:
```{r}
ggplot(MAG)+
  geom_line(aes(datetime, SO2))+
    scale_x_chron()
```

View raw particule matters 10 concentrations:
```{r}
ggplot(MAG)+
  geom_line(aes(datetime, PM10))+
    scale_x_chron()
```

View raw particule matters 2.5 concentrations:
```{r}
ggplot(MAG)+
  geom_line(aes(datetime, PM2.5))+
    scale_x_chron()
```

View raw EC concentrations:
```{r}
ggplot(MAG)+
  geom_line(aes(datetime, EC))+
    scale_x_chron()
```

View raw NOX concentrations:
```{r}
ggplot(MAG)+
  geom_line(aes(datetime, NOX))+
    scale_x_chron()
```

View raw temperatures:
```{r}
ggplot(MAG)+
  geom_line(aes(datetime, TEMP))+
    scale_x_chron()
```

View raw precipîtations:
```{r}
ggplot(MAG)+
  geom_line(aes(datetime, PREC))+
    scale_x_chron()
```

View raw global radiation:
```{r}
ggplot(MAG)+
  geom_line(aes(datetime, RAD))+
    scale_x_chron()
```

For Bern: 
View raw ozone concentrations:
```{r}
ggplot(BER)+
  geom_line(aes(datetime, O3))+
    scale_x_chron()
```

View raw nitrogen dioxide concentrations:
```{r}
ggplot(BER)+
  geom_line(aes(datetime, NO2))+
    scale_x_chron()
```

View raw particule matters 10 concentrations:
```{r}
ggplot(BER)+
  geom_line(aes(datetime, PM10))+
    scale_x_chron()
```

View raw particule matters 2.5 concentrations:
```{r}
ggplot(BER)+
  geom_line(aes(datetime, PM2.5))+
    scale_x_chron()
```

View raw EC concentrations:
```{r}
ggplot(BER)+
  geom_line(aes(datetime, EC))+
    scale_x_chron()
```

View raw NOX concentrations:
```{r}
ggplot(BER)+
  geom_line(aes(datetime, NOX))+
    scale_x_chron()
```

View raw temperatures:
```{r}
ggplot(BER)+
  geom_line(aes(datetime, TEMP))+
    scale_x_chron()
```

View raw precipitations:
```{r}
ggplot(BER)+
  geom_line(aes(datetime, PREC))+
    scale_x_chron()
```

View raw global radiation:
```{r}
ggplot(BER)+
  geom_line(aes(datetime, RAD))+
    scale_x_chron()
```


Create dataframes for comparison, only keeping the common colums. Copy dataframes and delete the desired columns:
```{r}
MAG1 <- MAG 
BER1 <- BER 

MAG1[ ,c('SO2')] <- list(NULL)
BER1[ ,c('CO', 'CPC')] <- list(NULL)
```
```{r}
head(MAG1)
```
```{r}
head(BER1)
```

Comparison of pollutants in two localities:
```{r}
lf_MAG <- gather(MAG, variable, value, -c(datetime, month, date))
```

```{r}
ggplot(lf_MAG) +
  facet_grid(variable~., scale="free_y") +  
  geom_line(aes(datetime, value))+
  scale_x_chron()
```

```{r}
ggplot(lf_MAG, aes(month, value)) +
  facet_grid(variable~., scale="free_y") +
  geom_bar(stat="summary", fun="mean") +
  geom_errorbar(stat="summary",
                fun.min=min, #function(x) quantile(x, .25),
                fun.max=max, #function(x) quantile(x, .75))+
                width=0.1)
```

```{r}
lf_BER <- gather(BER, variable, value, -c(datetime, month, date))
```

```{r}
ggplot(lf_BER) +
  facet_grid(variable~., scale="free_y") +  
  geom_line(aes(datetime, value))+
  scale_x_chron()
```

```{r}
ggplot(lf_BER, aes(month, value)) +
  facet_grid(variable~., scale="free_y") +
  geom_bar(stat="summary", fun="mean") +
  geom_errorbar(stat="summary",
                fun.min=min, #function(x) quantile(x, .25),
                fun.max=max, #function(x) quantile(x, .75))+
                width=0.1)
```